### 2020-07-07 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
more model trials, updates/new data, prelim results

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
  # mutate(dnbr_qualitative = case_when(
  #   dnbr < 0.1 ~ "unburned",
  #   dnbr < 0.27 ~ "low",
  #   dnbr < 0.66 ~ "moderate",
  #   dnbr > 0.66 ~ "high")) %>% 
  # mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
  #        dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality))

@@

@@

  - Mixed confier forests
    + response variable: type conversion
    + response variable: 4 mortality classes (none, low, momderate, high)
    + response variable: 3 mortality classes (none/low, moderate, high)
    + response variable: 2 mortality classes (none/low/moderate, high)
    + response variable: rdnbr
  - Coulter pines 
    + response variable: rdnbr
  - mixed conifer forests and coulter pines combined
    + response variable: rdnbr

@@

### Sample Size for each mortality class
```{r}
mortality_count <- mcf %>% 
  group_by(mortality_qual) %>% 
  count()

@@

```
None: 251
Low: 166
mod: 160
high: 158

@@

ndvi_rdnbr_corr <- lm(rdnbr_2007~ndvi, data=mcf)
summary(ndvi_rdnbr_corr)

@@

### RF manual mortality: 4 classes (none, low, moderate, high)

Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable
```{r}

@@

rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
                    data = mcf, 
                    sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

@@

```
variable importance seems stable AND class accuracy is high and equal between classes!!

#### ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)

## PDP v. ALE

```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(mcf), ndvi, "0") #0=not type converted
b <- partialPlot(rfm, as.data.frame(mcf), ndvi, "1", col="red") #1=type converted
lines(a, col="green")

@@

#ale (see detailed notes in ALE Plots code chunk below; not working right now)
train_x <- mcf[-which(names(mcf) == "tc_mortality")]
ndvi_predictor_notTC <- Predictor$new(rfm, data = train_x, y = mcf$tc_mortality, type = "prob", class = "0")
ndvi_not <- plot(FeatureEffect$new(ndvi_predictor_notTC, feature = "ndvi")) + ggtitle("not TC")

#predict values for the other mortality classes for the same variable
ndvi_predictor_tc <- Predictor$new(rfm, data = train_x, type = "prob", class = "1")
sr_moderate <- plot(FeatureEffect$new(ndvi_predictor_tc, feature = "ndvi")) + ggtitle("TC")
```
so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero. only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities. So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.

>>>>

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
  mutate(dnbr_qualitative = case_when(
    dnbr_2007 < 0.1 ~ "unburned",
    dnbr_2007 < 0.27 ~ "low",
    dnbr_2007 < 0.66 ~ "moderate",
    dnbr_2007 > 0.66 ~ "high")) %>%
  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>%
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High"))) %>% 
  mutate(two_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(two_class_mortality = as.factor(two_class_mortality)) %>% 
  mutate(two_class_mortality= fct_relevel(two_class_mortality, 
                                     levels=c("Other", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 69 ~ "None",
    rdnbr_2007 < 315 ~ "Low",
    rdnbr_2007 < 640 ~ "Moderate",
    rdnbr_2007 > 640 ~ "High")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
    mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
                                     levels=c("Other", "High")))

@@

pico <- read_csv("coulter_env_modified.csv") %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 69 ~ "None",
    rdnbr_2007 < 315 ~ "Low",
    rdnbr_2007 < 640 ~ "Moderate",
    rdnbr_2007 > 640 ~ "High")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
  na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
  mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
                                     levels=c("Other", "High"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect_30m)/360))/2) %>% 
  rename_all(funs(str_replace(., "_30m", ""))) %>% 
  mutate(dnbr_qualitative = case_when(
    dnbr_2007 < 0.1 ~ "unburned",
    dnbr_2007 < 0.27 ~ "low",
    dnbr_2007 < 0.66 ~ "moderate",
    dnbr_2007 > 0.66 ~ "high")) %>%
  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned")))

  #document name is "modified" bc i manually removed all the unnesessary arc columns

@@

  - Mixed confier forests
    + response variable: type conversion
    + response variable: 4 mortality classes (none, low, momderate, high)
    + response variable: 3 mortality classes (none/low, moderate, high)
    + response variable: 2 mortality classes (none/low/moderate, high)
    + response variable: rdnbr (regression)
  - Coulter pines 
    + response variable: rdnbr (regression)
  - mixed conifer forests and coulter pines combined
    + response variable: rdnbr (regression)

@@

### Sample Size for each mortality class
```{r}
#mcf -------------------------------------------------------------------
mortality_count <- mcf %>% 
  group_by(mortality_qual) %>% 
  count()

@@

rdnbr_mcf <- mcf %>% 
  group_by(rdnbr_qualitative) %>% 
  count()

dnbr_mcf <- mcf %>% 
  group_by(dnbr_qualitative) %>% 
  count()

#coulter ------------------------------------------------------------------- 
rdnbr_pico <- pico %>% 
  group_by(rdnbr_qualitative) %>% 
  count()

dnbr_pico <- pico %>% 
  group_by(dnbr_qualitative) %>% 
  count()

# all -------------------------------------------------------------------
all_sample_size <- cbind(mortality_count, rdnbr_mcf, dnbr_mcf, rdnbr_pico, dnbr_pico)
```
Yah, dnbr is not a good index to use. Rdnbr is better.

### mortality v rdnbr 
```{r}
rdnbr_qual_n <- mcf %>% 
  group_by(rdnbr_qualitative) %>% 
  count() %>% 
  mutate(mort_type = "rdnbr") %>% 
  rename(mortality_qual = rdnbr_qualitative)

mortality_n <- mcf %>% 
  group_by(mortality_qual) %>% 
  count() %>% 
  mutate(mort_type = "manual")

mortality_count <- rbind(rdnbr_qual_n, mortality_n)

ggplot(mortality_count, aes(x = mortality_qual, y = n, fill = mort_type)) + geom_col(position = "dodge") +
  geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
  scale_fill_discrete(name = "How was mortality determined")

## type conversion threshold for rdnbr
```{r}
##plot -------------------------------------------------------------------------------

# all
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter()

#by prefire cover
ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
  geom_jitter() +
  facet_wrap(~prefire)

## use lm to get a rdnbr threshold for the type converter stands? ---------------------

threshold <- lm(rdnbr_2007 ~ tc_mortality, data = mcf)
summary(threshold) #r2=0.53

threshold1 <- lm(rdnbr_2007 ~ tc_mortality + ndvi, data = mcf)
summary(threshold1) #r2=0.54

threshold2 <- lm(rdnbr_2007 ~ tc_mortality + prefire, data = mcf)
summary(threshold2) #r2=0.54

# max type converted values ----------------------------------------------

max_nonTC <- mcf %>% 
  group_by(tc_mortality) %>% 
  summarise(max = max(rdnbr_2007),
            min = min(rdnbr_2007)) 
  #non-type converted=1160.433 ; type converted=1414.205

# potential type conversion threshold --------------------------------------------

tc_threshold <- mcf %>% 
  mutate(rdnbr_tc_threshold = case_when(
    rdnbr_2007 <1160 ~ "not_tc",
    rdnbr_2007 >1160 ~ "tc"
  ))

```
well there some overlap with the the 2 classes, but there is also a clear point that the non type converted stands reach (rdnbr=1160) that many type converted stands surpass.

##high mortality v rdnbr
```{r}
rdnbr_qual_n <- mcf %>% 
  group_by(rdnbr_qualitative2) %>% 
  count() %>% 
  mutate(mort_type = "rdnbr") %>% 
  rename(two_class_mortality = rdnbr_qualitative2)

mortality_n <- mcf %>% 
  group_by(two_class_mortality) %>% 
  count() %>% 
  mutate(mort_type = "manual")

mortality_count <- rbind(rdnbr_qual_n, mortality_n)

ggplot(mortality_count, aes(x = two_class_mortality, y = n, fill = mort_type)) + geom_col(position = "dodge") +
  geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
  scale_fill_discrete(name = "How was mortality determined")

@@

ndvi_rdnbr_corr <- lm(rdnbr_2007~ndvi, data=mcf)
summary(ndvi_rdnbr_corr) #3%
ndmi_rdnbr_corr <- lm(rdnbr_2007~ndmi_2007, data=mcf)
summary(ndmi_rdnbr_corr) #82.7%

@@

### RF manual mortality: 4 classes (none, low, moderate, high)
```{r}
for (i in 1:10) {

rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('None' = 158, 'Low' = 158, 'Moderate' = 158, 'High' = 158), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```

PDP
```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "None") 
c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Moderate") 
b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
lines(a, c, d, col= "green", "yellow", "orange") # eh, not working. forgot how to set this up 2 show all 4 off the top of my head

a <- partialPlot(rfm, as.data.frame(mcf), slope, "None") 
c <- partialPlot(rfm, as.data.frame(mcf), slope, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), slope, "Moderate") 
b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8)) 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf), sr, "None") 
c <- partialPlot(rfm, as.data.frame(mcf), sr, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), sr, "Moderate")  
b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red") 
lines(a, col="green")

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(mortality_qual, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "mortality_qual")]

# predictors 
none_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="None")
low_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="Low")
moderate_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="Moderate")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="High")

#vpd_jan
none <- plot(FeatureEffect$new(none_predictor, feature = "vpd_jan")) + ggtitle("None")
low <- plot(FeatureEffect$new(low_predictor, feature = "vpd_jan")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "vpd_jan")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
ggarrange(none, low, moderate, high)

#slope
none <- plot(FeatureEffect$new(none_predictor, feature = "slope")) + ggtitle("None")
low <- plot(FeatureEffect$new(low_predictor, feature = "slope")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "slope")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
ggarrange(none, low, moderate, high)

#sr
none <- plot(FeatureEffect$new(none_predictor, feature = "sr")) + ggtitle("None")
low <- plot(FeatureEffect$new(low_predictor, feature = "sr")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "sr")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
ggarrange(none, low, moderate, high)

### RF manual mortality: 3 classes (none/low, moderate, high)
```{r}
for (i in 1:10) {

rfm <- randomForest(three_class_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('Low' = 158, 'Moderate' = 158, 'High' = 158),
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```

PDP
```{r}
#PDP
c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Moderate") 
b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8))

c <- partialPlot(rfm, as.data.frame(mcf), slope, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), slope, "Moderate") 
b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))

c <- partialPlot(rfm, as.data.frame(mcf), sr, "Low") 
d <- partialPlot(rfm, as.data.frame(mcf), sr, "Moderate")  
b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(three_class_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "three_class_mortality")]

# predictors 
low_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="Low")
moderate_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="Moderate")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="High")

#vpd_jan
low <- plot(FeatureEffect$new(low_predictor, feature = "vpd_jan")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "vpd_jan")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
ggarrange(low, moderate, high)

#slope
low <- plot(FeatureEffect$new(low_predictor, feature = "slope")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "slope")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
ggarrange(low, moderate, high)

#sr
low <- plot(FeatureEffect$new(low_predictor, feature = "sr")) + ggtitle("Low")
moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "sr")) + ggtitle("Moderate")
high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
ggarrange(low, moderate, high)

### RF manual mortality: 2 classes (none/low/moderate (ie other), high)
```{r}
for (i in 1:10) {

rfm <- randomForest(two_class_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('Other' = 158,'High' = 158),
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```
ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf$two_class_mortality, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)
```
PDP
```{r}
#PDP
c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
lines(c, col="green")

c <- partialPlot(rfm, as.data.frame(mcf), slope, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))

c <- partialPlot(rfm, as.data.frame(mcf), sr, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")

### RF manual mortality: 2 classes (none/low (ie other), high)
removed moderate class

mcf_2_class_mod <- mcf %>% 
  filter(mortality_qual != "Moderate") %>% 
  mutate(two_class_mortality_mod = case_when(
    mortality_qual %in% c("None", "Low") ~ "Other",
    mortality_qual %in% c("High") ~ "High"
  ))
mcf_2_class_mod$two_class_mortality_mod <- factor(mcf_2_class_mod$two_class_mortality_mod)  # RF still recognizes that there were 4 mortality classes even though filtered one of them out. This allows rf to not recognize the missing class.

@@

rfm <- randomForest(two_class_mortality_mod ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf_2_class_mod, 
                    sampsize = c('Other' = 158,'High' = 158),
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```

ROC/AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf_2_class_mod$two_class_mortality_mod, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)
```

PDP
```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), vpd_jan, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), slope, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), slope, "High", col="red", ylim=c(-0.8,0.8)) 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), sr, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), sr, "High", col="red") 
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), TPI_five, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), TPI_five, "High", col="red", ylim=c(-0.8,0.8)) 
lines(a, col="green")

ALE
```{r}
#sub data
mcf_ale <- mcf_2_class_mod %>% 
  select(two_class_mortality_mod, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "two_class_mortality_mod")]

# predictors 
other_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$two_class_mortality_mod, type = "prob", class="Other")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$two_class_mortality_mod, type = "prob", class="High")

#vpd_jan
other <- plot(FeatureEffect$new(other_predictor, feature = "vpd_jan")) + ggtitle("Other")
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
ggarrange(other, high)

#slope
other <- plot(FeatureEffect$new(other_predictor, feature = "slope")) + ggtitle("Other")
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
ggarrange(other, high)

#sr
other <- plot(FeatureEffect$new(other_predictor, feature = "sr")) + ggtitle("Other")
high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
ggarrange(other, high)

### RF manual: type conversion
Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable

0 = not type converted
1 = type converted
```{r}

rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

@@

```
variable importance is KINDA/MOSTLY stable. There arent big changes between runs, some variables flip 1 space, but some move 3 spaces. Variable importance was more stable when ndvi was included in the model.

ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)
```

PDP
```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "0") #0=not type converted
b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "1", col="red") #1=type converted
lines(a, col="green")

@@

```
so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero. only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities. So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(tc_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "tc_mortality")]

# predictors 
nottc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="0")
tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="1")

#vpd_jan
other <- plot(FeatureEffect$new(nottc_predictor, feature = "vpd_jan")) + ggtitle("Not Type Converted")
high <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_jan")) + ggtitle("Type Converted")
ggarrange(other, high)

#slope
other <- plot(FeatureEffect$new(nottc_predictor, feature = "slope")) + ggtitle("Not Type Converted")
high <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted")
ggarrange(other, high)

#sr
other <- plot(FeatureEffect$new(nottc_predictor, feature = "sr")) + ggtitle("Not Type Converted")
high <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted")
ggarrange(other, high)

# rdnbr: classification (2 classes)
```{r}
for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                    data = mcf, 
                    sampsize = c('High' = 150, 'Other' = 150), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```
get very similar oob and class accuracy as the manual mortality trial with 2 classes (none/low/moderate, high). So if we just look at high mortality v other mortality then we can use either rdnbr or manual mortality but if we chose to show the other classes it makes sense to use manual mortality since it is a lot more accurate, and thus more representative, than rdnbr.

ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf$rdnbr_qualitative2, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)
```

PDP
```{r}
#PDP
c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
lines(c, col="green")

c <- partialPlot(rfm, as.data.frame(mcf), slope, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))

c <- partialPlot(rfm, as.data.frame(mcf), sr, "Other") 
b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "rdnbr_qualitative2")]

# predictors 
other_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative2, type = "prob", class="Other")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative2, type = "prob", class="High")

#vpd_jan
other <- plot(FeatureEffect$new(other_predictor, feature = "vpd_jan")) + ggtitle("Other")
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
ggarrange(other, high)

#slope
other <- plot(FeatureEffect$new(other_predictor, feature = "slope")) + ggtitle("Other")
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
ggarrange(other, high)

#sr
other <- plot(FeatureEffect$new(other_predictor, feature = "sr")) + ggtitle("Not Type Converted")
high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("Type Converted")
ggarrange(other, high)

# rdnbr: regression
```{r}
for (i in 1:10) {

rfm <- randomForest(rdnbr_2007 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan, 
                    data = mcf, 
                    #sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```

PDP
```{r}
#PDP
partialPlot(rfm, as.data.frame(mcf), vpd_jan) 
partialPlot(rfm, as.data.frame(mcf), slope) 
partialPlot(rfm, as.data.frame(mcf), sr)

```
not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.

train_x <- mcf_ale[-which(names(mcf_ale) == "rdnbr_qualitative")]

# predictors 
predictor <- Predictor$new(rfm, data = mcf_ale)

#explanatory variables
vpd_jan <- plot(FeatureEffect$new(predictor, feature = "vpd_jan")) + ggtitle("vpd_jan")
slope <- plot(FeatureEffect$new(predictor, feature = "slope")) + ggtitle("slope")
sr <- plot(FeatureEffect$new(predictor, feature = "sr")) + ggtitle("sr")
ggarrange(vpd_jan, slope, sr)

```
not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)

# rdnbr: regression PICO
```{r}
for (i in 1:10) {

rfm <- randomForest(rdnbr_2007 ~ rain + tmax + TPI_five + dem  + insolation + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, #missing fow accumulation
                    data = pico, 
                    #sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
```

PDP
```{r}
#PDP
partialPlot(rfm, as.data.frame(pico), rain) 
partialPlot(rfm, as.data.frame(pico), slope) 
partialPlot(rfm, as.data.frame(pico), insolation)

```
not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)

ALE
```{r}
#sub data
mcf_ale <- mcf %>% 
  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
